---
title: "prototype"
author: "Stefanie Mueller"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    lightbox: true
    gallery: true
    toc_depth: 3
    use_bookdown: true
    fig_caption: true
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r knitr_init, echo=FALSE, cache=FALSE}
# load libraries
library(knitr)
library(rmdformats)

# define functions
  myDT = function(data){
  data %>% DT::datatable(rownames = FALSE,
    filter = 'bottom',
    extensions = 'Buttons',
      options = list(
        searching = TRUE,
        fixedColumns = TRUE,
        autoWidth = TRUE,
        dom = 'Bfrtip',
        buttons = c('copy', 'csv', 'excel'))
    )
}

## Global options
#options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=FALSE,
               prompt=TRUE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               fig.width = 10)
opts_knit$set(width=75)
```

# PREFACE

## PROJECT
Explore influence of political system on policy stringency and severity of 
COVID epidemic in individual countries.

## OBJECTIVE
protoype analysis between stringency, COVID infections and deaths and political
system

## load libraries
```{r}
library(tidyverse)
library(data.table)
library(maps)
library(viridis)
library(readxl)
library(ggpmisc)
```

## Version Info
```{r}
sessionInfo()
```

## Set Global Varibales
```{r}
DIR="/home/rstudio/NONWORK/HackathonUCL_2020-1/rstuff/"
OUTFOLDER=paste0(DIR,"/data/output/")
INPUTFOLDER=paste0(DIR,"/data/input/")
SCRIPT="S01"
DATE=Sys.Date()
```

# Load and Format Oxford Government Response data

Data Source:
[https://github.com/OxCGRT/covid-policy-tracker](https://github.com/OxCGRT/covid-policy-tracker)


## Load latest data version from github
```{r}
datain = fread("https://raw.githubusercontent.com/OxCGRT/covid-policy-tracker/master/data/OxCGRT_latest.csv")
```

## Save local copy
```{r}
write_csv(datain,
          paste0(OUTFOLDER, SCRIPT, ".oxford_government_reponse_latest.", DATE, ".csv.gz"))
```

### Glimpse at initial data, look at 10 random data rows
```{r}
datain %>% 
  sample_n(size=10) %>% myDT()
```


## Select Nationwide data and drop regional data
```{r}
data_nationwide = datain %>% 
  filter(Jurisdiction == "NAT_TOTAL") 
```

## Select data available till 1st of June 2021
```{r}
data_natiowide_june_1st = data_nationwide %>% 
  filter(Date=="20210601") 
```

# Make first map
## Total death COVID numbers till 1st of June 2021
```{r}
world_map <- map_data("world")
# rename some world mape countries for merging

world_map[world_map$region=="USA",]$region ="United States"
world_map[world_map$region=="UK",]$region ="United Kingdom"


plot_data01 = world_map %>% 
  left_join(data_natiowide_june_1st[,c("CountryName", "ConfirmedDeaths")], 
            by=c("region"="CountryName"))


ggplot(plot_data01, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = ConfirmedDeaths ), color = "white")+
  scale_fill_viridis_c(option = "C") +
  theme_void() 
```

## Policy Stringency Index 1st of June 2021
```{r}
plot_data02 = world_map %>% 
  left_join(data_natiowide_june_1st[,c("CountryName", "StringencyIndex")], 
            by=c("region"="CountryName"))

ggplot(plot_data02, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = StringencyIndex ), color = "white")+
  scale_fill_viridis_c(option = "C") +
  theme_void()
```


# Load further annotation data

Sources:   
- population numbers for 2020: [https://www.worldometers.info/world-population/population-by-country/](https://www.worldometers.info/world-population/population-by-country/)
- democracy index and government type: [https://en.wikipedia.org/wiki/Democracy_Index](https://en.wikipedia.org/wiki/Democracy_Index) scrapped [here](https://docs.google.com/spreadsheets/d/1d0noZrwAWxNBTDSfDgG06_aLGWUz4R6fgDhRaUZbDzE/edit?usp=sharing) 

## read democracy index
```{r}
democracy_index = readxl::read_xlsx(paste0(INPUTFOLDER,"_EIU-Democracy Indices - Dataset - v4.xlsx"), sheet=4) %>% 
  #only keep data for year 2020
  filter(time==2020) %>% 
  mutate(DemocracyIndex2020 = `Democracy index (EIU)`/10) %>% 
  select(CountryName = name,
        DemocracyIndex2020) %>% 
  mutate(Government_Style  =case_when(
    DemocracyIndex2020 >8 ~ "Full democracies",
    DemocracyIndex2020 >6 ~ "Flawed democracies",
    DemocracyIndex2020 >4 ~ "Hybrid regimes",
    DemocracyIndex2020>0 | DemocracyIndex2020==0 ~"Authoritarian regimes"))

# renaming some countries to merge data
democracy_index[democracy_index$CountryName=="Congo, Dem. Rep.",]$CountryName  = "Democratic Republic of Congo"
democracy_index[democracy_index$CountryName=="Congo, Rep.",]$CountryName = "Congo"
democracy_index[democracy_index$CountryName=="Hong Kong, China",]$CountryName = "Hong Kong"
democracy_index[democracy_index$CountryName=="United States",]$CountryName = "United States"

```

## world population
```{r}
population = readxl::read_xlsx(paste0(INPUTFOLDER,"_EIU-Democracy Indices - Dataset - v4.xlsx"), sheet=16) %>% 
  #only keep data for year 2020
  filter(time==2020) %>% 
  select(CountryName = name,
        PopulationNumber2020 = "Population") 

# renaming some countries to merge data
population[population$CountryName=="Congo, Dem. Rep.",]$CountryName  = "Democratic Republic of Congo"
population[population$CountryName=="Congo, Rep.",]$CountryName = "Congo"
population[population$CountryName=="Hong Kong, China",]$CountryName = "Hong Kong"
```


## merge annotation data
```{r}
data_natiowide_june_1st_annotation = data_natiowide_june_1st %>% 
  left_join(democracy_index) %>% 
  left_join(population) %>% 
  # derive further annotations
  mutate(TotalCases = ConfirmedCases,
         TotalDeaths = ConfirmedDeaths) %>% 
  mutate(Cases_per_100k = TotalCases/PopulationNumber2020 * 100000,
         Deaths_per_100k = TotalDeaths/PopulationNumber2020 * 100000)
  
```


# Analysis Influence of Governement System on Policy Stringency

## Plot Government System for chosen countries
```{r}
plot_data03 = world_map %>% 
  left_join(data_natiowide_june_1st_annotation[,c("CountryName", "Government_Style", "StringencyIndex", "DemocracyIndex2020")], 
            by=c("region"="CountryName"))

ggplot(plot_data03, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = Government_Style ), color = "white")+
  theme_void() 
```

## Compare Policy Stringency on Governement System
```{r}
ggplot(data_natiowide_june_1st_annotation, aes(x=Government_Style, y=StringencyIndex,  color=Government_Style)) +
   theme_bw() +
  #geom_point(position=position_jitter(width=0.3), alpha=0.6) +
  geom_boxplot(size=1.4,fill=NA, outlier.size=0) +
  geom_boxplot(size=1.4,fill=NA, outlier.size=0, color=alpha("black",0.3)) +
 
    # change titles here:
  labs(y="COVID-19 Policy Stringency", x="Governmental System",
       title="Differences in Policy by Government System") +
  theme(title = element_text(size=rel(1.3), face = "bold"),
        axis.title.y = element_text(vjust=1.5),
        axis.title.x = element_text(vjust=-1.5),
        axis.text = element_text(size=rel(1.25)),
        axis.text.x = element_text(angle = 30,hjust = 1),
        legend.text = element_text(size=rel(1.3)),
        strip.text = element_text(size=rel(1.3)),
        plot.margin = unit(c(1,1,1,2), "cm"),
        panel.grid.major = element_line(colour="grey60"))

```

## Association between Policy Stringency and Governement Index
```{r}
ggplot(data_natiowide_june_1st_annotation,
       aes(x=DemocracyIndex2020, y=StringencyIndex)) +
   theme_bw() +
  geom_point(aes(color=Government_Style)) +
  geom_smooth(method='lm', formula= y~x, se = FALSE) +
  stat_poly_eq(formula = y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) +
  theme(title = element_text(size=rel(1.3), face = "bold"),
        axis.title.y = element_text(vjust=1.5),
        axis.title.x = element_text(vjust=-1.5),
        axis.text = element_text(size=rel(1.25)),
        axis.text.x = element_text(angle = 30,hjust = 1),
        legend.text = element_text(size=rel(1.3)),
        strip.text = element_text(size=rel(1.3)),
        plot.margin = unit(c(1,1,1,2), "cm"),
        panel.grid.major = element_line(colour="grey60"))

```

## Association between Policy Stringency and Death_per_100K
```{r}
ggplot(data_natiowide_june_1st_annotation,
       aes(x=log10(Deaths_per_100k), y=StringencyIndex)) +
   theme_bw() +
  geom_point(aes(color=Government_Style)) +
  geom_smooth(method='lm', formula= y~x, se = FALSE) +
  stat_poly_eq(formula = y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) +
  theme(title = element_text(size=rel(1.3), face = "bold"),
        axis.title.y = element_text(vjust=1.5),
        axis.title.x = element_text(vjust=-1.5),
        axis.text = element_text(size=rel(1.25)),
        axis.text.x = element_text(angle = 30,hjust = 1),
        legend.text = element_text(size=rel(1.3)),
        strip.text = element_text(size=rel(1.3)),
        plot.margin = unit(c(1,1,1,2), "cm"),
        panel.grid.major = element_line(colour="grey60"))

```


## Assocaition between Policy Stringency and Cases_per_100K
```{r}
ggplot(data_natiowide_june_1st_annotation,
       aes(x=log10(Cases_per_100k), y=StringencyIndex)) +
   theme_bw() +
  geom_point(aes(color=Government_Style)) +
  geom_smooth(method='lm', formula= y~x, se = FALSE) +
  stat_poly_eq(formula = y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) +
  theme(title = element_text(size=rel(1.3), face = "bold"),
        axis.title.y = element_text(vjust=1.5),
        axis.title.x = element_text(vjust=-1.5),
        axis.text = element_text(size=rel(1.25)),
        axis.text.x = element_text(angle = 30,hjust = 1),
        legend.text = element_text(size=rel(1.3)),
        strip.text = element_text(size=rel(1.3)),
        plot.margin = unit(c(1,1,1,2), "cm"),
        panel.grid.major = element_line(colour="grey60"))

```


## Assocaition between DemocracyIndex in 2020 and Death_per_100K
```{r}
ggplot(data_natiowide_june_1st_annotation,
       aes(x=log10(Deaths_per_100k), y=DemocracyIndex2020)) +
   theme_bw() +
  geom_point(aes(color=Government_Style)) +
  geom_smooth(method='lm', formula= y~x, se = FALSE) +
  stat_poly_eq(formula = y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) +
  theme(title = element_text(size=rel(1.3), face = "bold"),
        axis.title.y = element_text(vjust=1.5),
        axis.title.x = element_text(vjust=-1.5),
        axis.text = element_text(size=rel(1.25)),
        axis.text.x = element_text(angle = 30,hjust = 1),
        legend.text = element_text(size=rel(1.3)),
        strip.text = element_text(size=rel(1.3)),
        plot.margin = unit(c(1,1,1,2), "cm"),
        panel.grid.major = element_line(colour="grey60"))

```

## Association between DemocracyIndex in 2020 and Cases_per_100K
```{r}
ggplot(data_natiowide_june_1st_annotation,
       aes(x=log10(Cases_per_100k), y=DemocracyIndex2020)) +
   theme_bw() +
  geom_point(aes(color=Government_Style)) +
  geom_smooth(method='lm', formula= y~x, se = FALSE) +
  stat_poly_eq(formula = y~x, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) +
  theme(title = element_text(size=rel(1.3), face = "bold"),
        axis.title.y = element_text(vjust=1.5),
        axis.title.x = element_text(vjust=-1.5),
        axis.text = element_text(size=rel(1.25)),
        axis.text.x = element_text(angle = 30,hjust = 1),
        legend.text = element_text(size=rel(1.3)),
        strip.text = element_text(size=rel(1.3)),
        plot.margin = unit(c(1,1,1,2), "cm"),
        panel.grid.major = element_line(colour="grey60"))

```


# Output Data for monogame
```{r}
output = data_natiowide_june_1st_annotation %>% 
  select("CountryName","PopulationNumber2020","Government_Style",
         "DemocracyIndex2020","TotalCases","TotalDeaths","Cases_per_100k","Deaths_per_100k","StringencyIndex")

```

## save
```{r}
write_csv(output,
          paste0(OUTFOLDER, SCRIPT, ".number_for_monogame.", DATE, ".csv"))
```

