---
title: "prototype"
author: "Stefanie Mueller"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    lightbox: true
    gallery: true
    toc_depth: 3
    use_bookdown: true
    fig_caption: true
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r knitr_init, echo=FALSE, cache=FALSE}
# load libraries
library(knitr)
library(rmdformats)

# define functions
  myDT = function(data){
  data %>% DT::datatable(rownames = FALSE,
    filter = 'bottom',
    extensions = 'Buttons',
      options = list(
        searching = TRUE,
        fixedColumns = TRUE,
        autoWidth = TRUE,
        dom = 'Bfrtip',
        buttons = c('copy', 'csv', 'excel'))
    )
}

## Global options
#options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=FALSE,
               prompt=TRUE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               fig.width = 10)
opts_knit$set(width=75)
```

# PREFACE

## PROJECT
Explore influence of political system on policy stringency and severity of 
COVID epidemic in individual countries.

## OBJECTIVE
protoype analysis between stringency, COVID infections and deaths and political
system

## load libraries
```{r}
library(tidyverse)
library(data.table)
library(maps)
library(viridis)
```

## Version Info
```{r}
sessionInfo()
```

## Set Global Varibales
```{r}
DIR="/home/rstudio/NONWORK/HackathonUCL_2020-1/rstuff/"
OUTFOLDER=paste0(DIR,"/data/output/")
INPUTFOLDER=paste0(DIR,"/data/input/")
SCRIPT="S01"
DATE=Sys.Date()
```

# Load and Format Oxford Government Response data
Data Source:
[https://github.com/OxCGRT/covid-policy-tracker](https://github.com/OxCGRT/covid-policy-tracker)


## Load latest data version from github
```{r}
datain = fread("https://raw.githubusercontent.com/OxCGRT/covid-policy-tracker/master/data/OxCGRT_latest.csv")
```

## Save local copy
```{r}
write_csv(datain,
          paste0(OUTFOLDER, SCRIPT, ".oxford_government_reponse_latest.", DATE, ".csv.gz"))
```

### Glimpse at initial data, look at 10 random data rows
```{r}
datain %>% 
  sample_n(size=10) %>% myDT()
```


## Select Nationwide data and drop regional data
```{r}
data_nationwide = datain %>% 
  filter(Jurisdiction == "NAT_TOTAL") 
```

## Select data available till 1st of June 2021
```{r}
data_natiowide_june_1st = data_nationwide %>% 
  filter(Date=="20210601") 
```

# Make first map
## Total death COVID numbers till 1st of June 2021
```{r}
world_map <- map_data("world")
plot_data01 = world_map %>% 
  left_join(data_natiowide_june_1st[,c("CountryName", "ConfirmedDeaths")], 
            by=c("region"="CountryName"))

ggplot(plot_data01, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = ConfirmedDeaths ), color = "white")+
  scale_fill_viridis_c(option = "C") +
  theme_void() 
```

## Policy Stringency Index 1st of June 2021
```{r}
plot_data02 = world_map %>% 
  left_join(data_natiowide_june_1st[,c("CountryName", "StringencyIndex")], 
            by=c("region"="CountryName"))

ggplot(plot_data02, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = StringencyIndex ), color = "white")+
  scale_fill_viridis_c(option = "C") +
  theme_void() 
```


# Load further annotation data
Sources:
- population numbers for 2020: [https://www.worldometers.info/world-population/population-by-country/](https://www.worldometers.info/world-population/population-by-country/)
- democracy index and government type: [https://en.wikipedia.org/wiki/Democracy_Index](https://en.wikipedia.org/wiki/Democracy_Index)
```{r}
annotation = fread(paste0(INPUTFOLDER, "further_annotations.csv"))
```

## look at annotations for some examplary countries
```{r}
annotation %>% myDT()
```

## Combine with Oxford data
```{r}
data_natiowide_june_1st_annotation = data_natiowide_june_1st %>% 
  inner_join(annotation, by="CountryName")
```

# Analysis Influence of Governement System on Policy Stringency

## Plot Government System for chosen countries
```{r}
plot_data03 = world_map %>% 
  left_join(data_natiowide_june_1st_annotation[,c("CountryName", "Government_Style", "StringencyIndex")], 
            by=c("region"="CountryName"))

ggplot(plot_data03, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = Government_Style ), color = "white")+
  theme_void() 
```

## Compare Policy Stringency
```{r}
ggplot(plot_data03, aes(x=Government_Style, y=StringencyIndex,
                        color=Government_Style)) +
   theme_bw() +
  #geom_point(position=position_jitter(width=0.3), alpha=0.6) +
  geom_boxplot(size=1.4,fill=NA, outlier.size=0) +
  geom_boxplot(size=1.4,fill=NA, outlier.size=0, color=alpha("black",0.3)) +
 
    # change titles here:
  #labs(y="Y Axis title", x="X Axis Title", title="Title") +
  theme(title = element_text(size=rel(1.3), face = "bold"),
        axis.title.y = element_text(vjust=1.5),
        axis.title.x = element_text(vjust=-1.5),
        axis.text = element_text(size=rel(1.25)),
        axis.text.x = element_text(angle = 30,hjust = 1),
        legend.text = element_text(size=rel(1.3)),
        strip.text = element_text(size=rel(1.3)),
        plot.margin = unit(c(1,1,1,2), "cm"),
        panel.grid.major = element_line(colour="grey60"))

```

